Bioverse: Potentially Observable Exoplanet Biosignature Patterns Under the UV Threshold Hypothesis for the Origin of Life

### 2025-02-05 src/tex/ms.tex
Introducing bayes factor in methods

\subsection{Semi-analytical approach}\label{sec:met-semianalytical}
%First, we apply a Bayes Factor Design Analysis~\citep{Schoenbrodt2018}
We first assessed the expected probabilities of obtaining true negative or true positive evidence for the UV Threshold Hypothesis ($H_1$) above, as well as the probability for misleading or inconclusive evidence, under idealized conditions.
This serves as a first-order estimate of the information content of a survey, before we take into account the effects of exoplanet demographics, sample selection, and survey strategy.

@@

@@

Let $Y=\sum_i^n L_i$ be the random variable counting the number of positive life detections in a sample of size $n$. Its probability mass function under the null hypothesis $H_\mathrm{null}$ is that of a binomial distribution:
\begin{equation}
P(Y=k|H_\mathrm{null}) = \binom{n}{k}f_\mathrm{life}^k(1-f_\mathrm{life})^{n-k}.
\end{equation}

>>>>

\subsection{Semi-analytical approach}\label{sec:met-semianalytical}
%First, we apply a Bayes Factor Design Analysis~\citep{Schoenbrodt2018}
\rev{In order to constrain the evidence yield of potential future surveys aimed at evaluating the} UV Threshold Hypothesis ($H_1$) above, \rev{we first derive a simplified analytical approach.
This approach ignores the effects of exoplanet demographics and sample selection to focus on the qualitative effects of sample size and the unknown probablity of abiogenesis on the evidence yield of future surveys.
To do so, we need to frame hypotheses $H_\mathrm{null}$ and $H_1$ as probabilistic models that determine the probability distributions of outcomes in terms of samples of exoplanets for which biosignatures are detected or not.}
%We first assessed the expected probabilities of obtaining true negative or true positive evidence for the UV Threshold Hypothesis ($H_1$) above, as well as the probability for misleading or inconclusive evidence, under idealized conditions.
%This serves as a first-order estimate of the information content of a survey, before we take into account the effects of exoplanet demographics, sample selection, and survey strategy.

@@

\rev{Next, we determine the probability distribution of sample outcomes, or likelihood of each hypothesis.}
Let $Y=\sum_i^n L_i$ be the random variable counting the number of positive life detections in a sample of size $n$. Its probability mass function under the null hypothesis $H_\mathrm{null}$ is that of a binomial distribution:
\begin{equation}
    \label{eq:semian:likelihoodHnull}
P(Y=k|H_\mathrm{null}) = \binom{n}{k}f_\mathrm{life}^k(1-f_\mathrm{life})^{n-k}.
\end{equation}

\rev{Then, we turn to a metric which measures the information gain of sampling: the Bayes factor \cite{Jeffreys1939}.
The posterior evidence for the null hypothesis $H_{\mathrm{null}}$ is given by the Bayes' theorem:
\begin{equation}
\label{eq:bayesHnull}
    P(H_{\mathrm{null}}|Y) = \frac{P(Y|H_{\mathrm{null}})P(H_{\mathrm{null}})}{P(Y)}
\end{equation}
and the posterior evidence for the NUV hypothesis:
\begin{equation}
    \label{eq:bayesHone}
        P(H_1|Y) = \frac{P(Y|H_1)P(H_1)}{P(Y)}~.
\end{equation}
Computing the distributions given by equations \ref{eq:bayesHnull} and \ref{eq:bayesHone} requires the distributions of likelihoods, given in equations \ref{eq:semian:likelihoodHnull} and \ref{eq:semian:likelihoodH1}, as well as the prior distributions for each hypothesis ($P(H_{\mathrm{null}})$ and $P(H_{1})$) and for the sample ($P(H_{1})$).
}
\rev{
Appart from the caveat that the likelihood for $H_{1}$ cannot be expressed analytically due to being conditioned by n_{\lambda} (equation \ref{eq:semian:likelihoodH1}), computing posterior probabilities is challenging due to the unknown priors.
}

@@
